# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
- histogram of the total number of steps taken each day
```{r echo=TRUE}
date_steps <- aggregate (steps ~ date, data = data, sum)
hist(date_steps$steps, main = "Steps taken per day", xlab = "Number of steps")
```
- mean and median total number of steps taken per day
```{r echo=TRUE}
mean_val<-round(mean(date_steps$steps),0)
median_val<-round(median(date_steps$steps),0)
```
 - `r paste("Mean:", mean_val)`
 - `r paste("Median:", median_val)`

## What is the average daily activity pattern?
```{r echo=TRUE}
intervals <- aggregate(steps ~ interval, data = data, mean)
plot(intervals$interval, intervals$steps, type="l", main = "Average daily activity pattern", xlab = "5-minute interval", ylab = "average number of steps taken")
max_step_inv <- intervals[which.max(intervals$steps) , "interval"]
```
- The 5-minute interval, on average across all days in the dataset, contains the maximum number of steps is **`r max_step_inv`**

## Imputing missing values  
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
```{r echo=TRUE}
missing <- length(data[complete.cases(data) == F,1])
```
- total number of missing values is **`r missing`**  

The new strategy for missing value is that **filling the mean steps for each interval**
```{r echo=TRUE}
mean_of_intervals <- function() {
  steps <- data.frame(levels(factor(data$interval)), 0)
  colnames(steps) <- c("interval", "steps")
  for(i in 1:length(steps[,1])) {
    m <- mean(data[data$interval == as.vector(steps[i,]$interval) & !is.na(data$steps), "steps"], na.rm = T)
    steps[i,]$steps <- round(m, 0)
  }
  steps
}
```
Use the new strategy to create a new data set
```{r echo=TRUE}
create_new_dataset <- function() {
  steps <- mean_of_intervals()
  new_data <- data
  for(i in 1:length(new_data[,1])) {
    if(is.na(new_data[i, 1])) {
      new_data[i, 1] <- steps[steps$interval == new_data[i, "interval"], 2]
    }
    if(weekdays(as.Date(new_data[i, "date"])) %in% c("Saturday","Sunday")) {
      new_data[i, "weekdays"] <- "weekend"
    }else {
      new_data[i, "weekdays"] <- "weekday"
    }
  }
  new_data
}
new_data <- create_new_dataset()
```

- histogram of the total number of steps taken each day based on new data set
```{r echo=TRUE}
date_steps <- aggregate(steps ~ date, data = new_data, sum)
hist(date_steps$steps, main = "Steps taken per day", xlab = "Number of steps")
```
- mean and median total number of steps taken per day based on new data set
```{r echo=TRUE}
mean_val<-round(mean(date_steps$steps),0)
median_val<-round(median(date_steps$steps),0)
```
 - `r paste("Mean:", mean_val)`
 - `r paste("Median:", median_val)`

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
library(lattice)
date_steps <- aggregate(steps ~ interval + weekdays, data = new_data, mean)
xyplot(steps ~ interval | weekdays, data = date_steps, layout = c(1, 2), type = "l", 
       xlab="5-minute interval", ylab="Number of Steps",
       main="Activity Patterns between Weekends and Weekdays")
```